
## Morphology

\paragraph{} Two key morphological features for distinguishing *Physaria*
sensu stricto include the membranous partition separating fruit locules
(i.e., replum; Figure \@ref(fig:FigureDiscussionReplumShape)) and the number of
ovules per locule (Figure \@ref(fig:FigureDiscussionOvules)). Replum shape
differs in species with a constricted replum noted as a linear appearance,
from those with a wider replum varying from oblong or elliptic to oblanceolate
or obovate. Species in the former group include the widespread *P. acutifolia*,
*P. vitulifera*, and *P. brassicoides*, while the latter group includes
*P. didymocarpa* and *P. saximontana* subspecies, *P. eburniflora*,
*P. integrifolia*, *P. dornii*, and *P. condensata*. The distinction of
partition shape and the quantity of ovules as observed by funiculi per locule
constitutes a primary feature in the dichotomous key of Wyoming *Physaria*
provided by @Dorn2001.

\paragraph{} Species consistently showing 2 ovules per locule include those with
linear partitions: *Physaria acutifolia*, *P. brassicoides*, and *P. vitulifera*.
@Rollins1984 described *P. saximontana* subsp. *saximontana* as having 2 ovules
per locule, differentiated from *P. s.* subsp. *dentata* with 2-4 ovules per
locule; this distinction is followed by @Dorn2001, @Lesica2012, and @OKane2010.
Material reviewed here indicates specimens from southeast of the Wind River
Range northwest through the Absaroka Range and Rocky Mountains of Montana
frequently have more than 2 locules per ovule. *P. eburniflora* of Natrona and
Carbon counties in Wyoming also frequently have 2-4 ovules per locule,
consistent with the description by @Rollins1981. Widespread *P. didymocarpa*
subsp. *didymocarpa* joins the southwestern Wyoming endemics *P. integrifolia*,
*P. dornii*, and *P. condensata* as consistently having more than 2 ovules per
locule. Notably, *P. didymocarpa* subsp. *lanata* in Big Horn, Sheridan, and
Johnson counties of Wyoming has a typically oblong replum with only 2 ovules
per locule. The main character used by @Rollins1984 to differentiate
*Physaria saximontana* from *P. didymocarpa* is the lack of a basal sinus in
the mature siliques of the former. This traditional circumscription is followed
by @Lesica2012 for Montana *Physaria* and @HeidelOKane2012 for Wyoming
*Physaria*. While *P. didymocarpa* and *P. eburniflora* are described as having
cordate basal sinuses [@Rollins1984; @Rollins1981], lack of a basal sinus
is observed throughout Montana *P. didymocarpa* specimens and seen in some
*P. eburniflora* individuals (Figure \@ref(fig:FigureDiscussionFruitApices)).
In contrast, *P. integrifolia*, *P. dornii*, and *P. condensata* from Teton,
Lincoln, Sublette, and Uinta counties of Wyoming typically have approximately
equal sinuses similar to the widespread *P. acutifolia*.

\paragraph{} In describing *Physaria didymocarpa* var. *integrifolia*
Rollins (later, *Physaria integrifolia* (Rollins) @Lichvar1984), @Rollins1939
noted appressed stellate trichomes on the mature siliques and entire basal
leaves as distinguishing characters from *P. d.* subsp. *didymocarpa*.
These characters are shared by the southwestern endemics *P. dornii* and
*P. condensata*, both of which are differentiated from *P. integrifolia*
by their condensed rosette habit [@Lichvar1983]
(Figure \@ref(fig:FigureDiscussionFruitTrichomes),
Figure \@ref(fig:FigureDiscussionBasalLeafMargins)).
In describing *P. saximontana*, @Rollins1984 noted the silique trichomes with
rays spreading away from the valve, as similarly described for *P. didymocarpa*
subsp. *didymocarpa* [@Rollins1939]. These observations are consistent with
material reviewed here and in keeping with keys by @Dorn2001 and
@HeidelOKane2012. Of the two *P. saximontana* varieties (var. *saximontana* and
var. *dentata*) described by @Rollins1984, the former was distinguished by
entire radical leaves and a relatively shorter style length from the latter
having broadly dentate margins and a longer style. While this traditional
distinction was followed by @OKane2010 and @Lesica2012, dentate margins can be
observed throughout *P. saximontana* of Wyoming and *P. didymocarpa* of Montana
(Figure \@ref(fig:FigureDiscussionBasalLeafMargins)). Though common in occurrence,
dentate margins are not supported as a fixed trait in *P. didymocarpa*.
In contrast, *P. acutifolia* and the southwestern endemics *P. integrifolia*,
*P. dornii*, and *P. condensata* generally have entire basal leaves with the
former distinguished from the endemics by replum shape and ovule count.

```{r FigureDiscussionReplumShape, fig.cap=caption$traitReplumShape, fig.scap="Distribution of fruit replums"}

caption$traitReplumShape <-
  "Distribution of discrete replum shapes by specimen location."

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/DiscussionTraitReplums",
    ext = knit_variables$graphics
  )
)

```

```{r FigureDiscussionOvules, fig.cap=caption$traitOvules, fig.scap="Distribution of fruit ovules"}

caption$traitOvules <-
  "Maximum count of ovules per locule by specimen location."

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/DiscussionTraitOvules",
    ext = knit_variables$graphics
  )
)

```

```{r FigureDiscussionFruitApices, fig.cap=caption$traitFruitApices, fig.scap="Distribution of mature silique sinuses"}

caption$traitFruitApices <-
  "Observations of mature silique sinuses by specimen location."

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/DiscussionTraitApices",
    ext = knit_variables$graphics
  )
)

```

```{r FigureDiscussionFruitTrichomes, fig.cap=caption$traitFruitTrichomes, fig.scap="Distribution of silique trichomes"}

caption$traitFruitTrichomes <-
  "Observations of mature silique trichomes by specimen location."

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/DiscussionTraitTrichomes",
    ext = knit_variables$graphics
  )
)

```

```{r FigureDiscussionBasalLeafMargins, fig.cap=caption$traitBasalLeafMargins, fig.scap="Distribution of basal leaf margins"}

caption$traitBasalLeafMargins <-
  "Observations of basal leaf margins by specimen location."

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/DiscussionTraitMargins",
    ext = knit_variables$graphics
  )
)

```
