# INTRODUCTION {#INTRODUCTION}

{{< include /_setup.qmd >}}

\pagestyle{main}
\pagenumbering{arabic}

\paragraph{} This project considers the systematics of *Physaria*
(Nuttall ex Torrey \& A. Gray) A. Gray [Brassicaceae] sensu stricto with an
emphasis on the regional and state endemics of Wyoming and surrounding areas.
Historically, these species of *Physaria* were circumscribed from
close relatives on the basis of didymous fruits, in which two highly inflated
valves are separated by a constricted replum [@Gray1848; @Watson1888], rather
than uninflated, unlobed fruits where the width of the replum is the same as
that of the fruit. Twelve such species have described distributions at least
partially in Wyoming:

* ***Physaria acutifolia*** Rydberg
* ***Physaria brassicoides*** Rydberg
* ***Physaria condensata*** Rollins
* ***Physaria didymocarpa*** subsp. ***didymocarpa*** (Hooker) A. Gray
* ***Physaria didymocarpa*** subsp. ***lanata*** (C.L. Hitchcock) O’Kane & Al-Shehbaz
* ***Physaria didymocarpa*** subsp. ***lyrata*** C.L. Hitchcock
* ***Physaria dornii*** Lichvar
* ***Physaria eburniflora*** Rollins
* ***Physaria integrifolia*** (Rollins) Lichvar
* ***Physaria saximontana*** subsp. ***saximontana*** Rollins
* ***Physaria saximontana*** subsp. ***dentata*** Rollins
* ***Physaria vitulifera*** Rydberg

\paragraph{} Multiple authors have noted the difficulty in delineating between
species traditionally placed in *Physaria* and those placed in the closely related
genus *Lesquerella* Watson. As first described by @Watson1888, *Lesquerella*
comprised 33 species previously recognized as the Old World *Vesicaria*.
Watson distinguished *Lesquerella* by the presence of a veined septum and the
absence of teeth or appendages on the filaments. Notably, he supported the
recognition of *Physaria* as a distinct entity presumably on the basis of
fruit morphology. In a monograph of *Lesquerella*, @Payson1921 remarked of
*L. utahensis* Rydberg, “$\ldots$great similarity, in some of its forms particularly,
to members of the genus *Physaria*.” @Rollins1939 noted further similarity,
stating that, “These two genera have almost exactly the same floral pattern,
habit of growth, and trichome morphology. The siliques too are very similar.”
@AlShehbaz2002 merged the genera on the basis of molecular, morphological,
ecological and distributional data, encapsulating more than 150
years of botanical progress.

\paragraph{} Following the combination of *Lesquerella* with *Physaria*
sensu stricto and the ensuing treatment by @OKane2010, a need for additional
systematic work among the “traditional” *Physaria* of Wyoming was identified
by @HeidelOKane2012. In particular, the contested status of specimens
previously annotated as *P. vitulifera* from south-central Wyoming is of
interest as the species is known primarily from the eastern slope of the
Front Range in Colorado [@Weber1996; @Ackerfield2015].
Furthermore, @OKane2010 previously noted the characters distinguishing
*P. didymocarpa* and *P. saximontana* show some intergradation, suggesting the
need for investigation of these taxa and described subspecies.
Lastly, the state endemic *P. dornii* of southwestern Wyoming has a described
distribution overlapping with the regional endemic *P. integrifolia* and a
peripatric distribution with *P. condensata* [@Lichvar1983].
Here, an integrative taxonomic treatment of these species considers evidence
from molecular, ecological and morphological data to address the
phylogenetic status of these taxa.

## Systematic and Taxonomic History

\paragraph{} The first mention of *Physaria* (Nutt.) A. Gray as a distinct taxon
is credited to Thomas Nuttall in Torrey and Gray [-@Nuttal1838] where it is
described as a section of *Vesicaria* Adanson, an Old World genus. At that time,
only *V. didymocarpa* Hooker as recognized by William Hooker was included. In
the description, @Hooker1830 noted of *V. didymocarpa* fruit morphology,
“...a seed-vessel with its two valves separated by so deep a sinus, that, ...it
might almost form a generic character.”  However, due to the remarkable
similarity of the overall habit to other *Vesicaria* species, @Hooker1830
chose to recognize *V. didymocarpa* at the specific rank.

\paragraph{} @Hooker1847 later described *Vesicaria geyeri* Hooker stating it
resembled *V. didymocarpa* in habit but could be differentiated by compressed,
obcordate fruits of membranous texture. @Gray1848 posited that these species
be recognized at the generic rank on the basis of strongly didymous fruits with
narrowed partitions, thereby recognizing *V. didymocarpa* and *V. geyeri* as the
genus *Physaria* (Nutt.) A. Gray. Gray [@Ives1861] made the next
addition to the genus with a description of *P. newberryi* A. Gray.
He distinguished this species from *P. didymocarpa* (Hook.) A. Gray by
highly inflated silicles with inflexed, or v-shaped apical sinuses.

\paragraph{} @Watson1882 made the final addition to the genus before the turn
of the century with a description of *Physaria oregona* S. Watson. He noted the
fruit valves as somewhat compressed and barely surpassing the replum,
round–reniform, and larger than *P. geyeri* (Hook.) A. Gray. A summary of
*Physaria* by @Robinson1895 included the four species so far mentioned and noted
them as similar in habit to *Lesquerella* S. Watson, but with distinguishable
characters in the fruit. Through the twentieth century, workers on these genera
largely upheld the “traditional” distinction between *Lesquerella* and
*Physaria* by the strongly didymous fruits of the latter.

\paragraph{} Per Axel Rydberg [-@Rydberg1901] made three additions to the genus
by describing *Physaria vitulifera* Rydberg, *P. floribunda* Rydberg, and
*P. acutifolia* Rydberg. These species were distinguished from *P. didymocarpa*
as having smaller fruits. The lack of a proximal (basal) sinus, curved pedicels,
and pandurate basal leaves of 4–5 cm further differentiated *P. vitulifera*
from *P. didymocarpa*. *Physaria floribunda* was described as an intermediate
species between *P. didymocarpa* and *P. vitulifera* distinguished by taller
height, deeply sinuate to slightly fiddle-form leaves of 5–10 cm with sinuately
toothed to acute terminal lobes, and a shallow basal sinus. *P. acutifolia* was
separated by the presence of deep apical and shallow basal sinuses, spherical
valves, and acute basal leaves of 1.5–2.5 cm. @Rollins1939 misapplied the name
*P. acutifolia* to a population from Gunnison Co., Colorado later described as
*P. rollinsii* G. A. Mulligan by @Mulligan1966. This discrepancy was highlighted
by Mulligan following cytological investigations of *P. acutifolia*
[@Mulligan1966; @Mulligan1967b]. @Rollins1981 later agreed an earlier treatment
[@Rollins1939] of *P. australis* (Payson) Rollins actually pertained to
*P. acutifolia* Rydberg.

\paragraph{} @Rydberg1902 made another addition to the genus with a
description of *Physaria brassicoides* Rydberg. *P. brassicoides* was
differentiated from *P. didymocarpa* by smaller fruit with an acute base in
addition to larger basal leaves forming a rosette of 7–10 cm in diameter.
Additionally, it was noted the plants tend to grow in soils rich with magnesium.
@Rollins1939 highlighted the constricted replum, presence of two ovules per
locule, entire basal leaves, and near lack of a basal sinus in this species.
Aven Nelson [-@Nelson1904] made the next addition to the genus by describing
a new variety of *Physaria didymocarpa*. Nelson distinguished *P. didymocarpa*
(Hook.) A. Gray var. *lanata* A. Nelson by the often irregularly toothed basal
leaves with marked pubescence of simple trichomes approaching a tomentose
appearance, with a type specimen collected from the Big Horn Mountains.
@Rydberg1912 recommended the species be recognized at the specific rank, a
status rejected by both @Payson1918 and @Rollins1939. @OKane2007 later changed
the status of *P. didymocarpa* var. *lanata* to the subspecific rank recognized
as *P. didymocarpa* (Hook.) A. Gray subsp. *lanata* (A. Nelson) O’Kane.

\paragraph{} A species from south central Washington, *Physaria alpestris*
Suksdorf was recognized by @Suksdorf1906 on the basis of compressed, terete to
obcordate fruits with styles from 4–6 mm. In a summary of Idaho threatened and
endangered plants, @Steele1977 reports *P. alpestris* Suskdorf var. *lyrata*
and *P. alpestris* Suksdorf var. *purpurea* as two varieties proposed for
federal threatened status. There, *Physaria geyeri* (Hook.) A. Gray var.
*purpurea* Rollins was listed as a synonym for *P. alpestris* var. *purpurea*.
The synonym *P. didymocarpa* (Hook) A. Gray var. *lyrata* C. L. Hitchcock is
listed for *P. alpestris* var. *lyrata*, but no reference for the varietal name
is provided. The author cites a single collection from Lemhi county shown near
Salmon, Idaho, stating variety *lyrata* shows intergradation with
*P. didymocarpa* var. *didymocarpa* populations near Malad in Oneida County
of south-central Idaho as well as populations in north-central Montana. It is
unclear where the name *Physaria alpestris* var. *lyrata* originated, though it
is listed as a synonym in a later report on the status of *P. didymocarpa* var.
*lyrata* in the Salmon National Forest [@Moseley1990]. *P. didymocarpa* var.
*lyrata* was first described by @Hitchcock1964 as a separate variety due to the
generally lyrate basal leaves and petioles in addition to the highly inflated,
larger silicles of variety lyrata. @OKane2007 changed the status of this
variety to the subspecific rank recognized as *P. didymocarpa* (Hook.) A. Gray
subsp. *lyrata* (C. L. Hitchcock) O’Kane.

\paragraph{} @Rollins1939 made notable contributions to *Physaria* with a
monograph of the genus wherein a number of new species and varieties were
described. Of note to the Wyoming flora are *P. condensata* Rollins,
*P. australis* (Payson) Rollins, and varieties of *P. didymocarpa*.
*Physaria condensata* was described from southwest Wyoming and separated on the
basis of a highly reduced growth-form while sharing fruit characteristics with
*P. didymocarpa*. The stems of *P. condensata* are described as less than a
centimeter in length with the entire plant generally forming a tuft under 5cm
in diameter. Additionally, @Rollins1939 mentions individuals of this species
form either soboles or single caudices, an example of intraspecific variation
in caudex branching.

\paragraph{} @Rollins1939 applied the name *Physaria australis* to
*P. didymocarpa* (Hook.) A. Gray var. *australis* Payson, which had been
previously described by @Payson1918. He argued the taxon deserved specific
status on the basis of a reduction from four to two ovules per locule and a
narrowly constricted linear replum. As mentioned previously, the name
*P. australis* (Payson) Rollins pertains to *P. acutifolia* Rydberg as it is
currently recognized [@Rollins1981]. @Rollins1939 also described
*P. didymocarpa* (Hook.) A. Gray var. *integrifolia* Rollins as a new variety of
*P. didymocarpa*, separated by appressed fruit trichomes. Individuals of this
variety otherwise tend to show replum structure and ovule number similar to
*P. didymocarpa* (Hook.) A. Gray var. *normalis* Kuntze. Variety *integrifolia*
was observed as sharing transitional characters with its southern analogue,
*P. acutifolia*, where  *P. acutifolia* was hypothesized to have diverged more
recently.

\paragraph{} @Lichvar1984 elevated *Physaria didymocarpa* var. *integrifolia*
to specific status on the basis of fruit, leaf, and trichome morphology. As
noted by @Rollins1983, fruit valves of *P. didymocarpa* var. *integrifolia* are
generally thin and highly inflated. In addition to fruit morphology differences,
Lichvar separated *P. integrifolia* (Rollins) Lichvar from *P. didymocarpa*
varieties *didymocarpa* and *lanata* on the basis of entire basal leaves and
siliques with appressed trichomes in *P. integrifolia*. In contrast,
*P. didymocarpa* var. *didymocarpa* and var. *lanata* both are described as
having dentate basal leaves and siliques with spreading trichomes. Lichvar also
highlighted differences in scanning electron micrograph images that show a
central umbo in var. *integrifolia* that is lacking in both var. *didymocarpa*
and var. *lanata* [@Rollins1979].

\paragraph{} @Lichvar1984 also described a new variety, *P. integrifolia*
(Rollins) Lichvar var. *monticola* Lichvar, as differing from variety
*integrifolia* by the presence of a highly branched caudex. The inflorescence
of variety *monticola* was described as not exceeding the basal rosette to
the same extent as variety *integrifolia*.  Additionally, both leaf and petiole
length were described as reduced in variety *monticola* relative to variety
*integrifolia*. Specimens identified *a priori* as variety *monticola* are
distributed throughout the range of variety *integrifolia*. @OKane2010 noted
the trait of branched caudices is likely due to caudex elongating in response
to shifting substrate conditions and does not constitute a taxonomic
delineation. Branching caudices also occur in other species of *Physaria*,
including *P. acutifolia*, *P. brassicoides*, and *P. didymocarpa*.

\paragraph{} Nearly forty years after his first monograph, @Rollins1981
described a whitish-flowered species as *Physaria eburniflora* Rollins based
on specimens from central Wyoming. In addition to the petal color,
*P. eburniflora* was distinguished by ascending trichomes similar to
*P. didymocarpa* but with fruit valves less inflated and of a rougher texture.
*Physaria eburniflora* was separated from both *P. acutifolia* and
*P. didymocarpa* by the shallow, slightly cordate basal sinus of the former as
opposed to the nearly equal, deep basal and apical sinuses of the latter two.
The basal leaves of *P. eburniflora* were described as entire and nearly
orbicular and the stems of the plant as prostrate. Lastly, Rollins described
*P. eburniflora* by a somewhat reduced growth-form and habit more similar to
*P. condensata* than *P. didymocarpa*.

\paragraph{} Following fieldwork in Wyoming, @Lichvar1983 described
*Physaria dornii* Lichvar as one of three *Physaria* populations with plants
showing a condensed rosette habit. [@Rollins1939; @Rollins1981] described the
other two populations as *P. condensata* and *P. eburniflora*. To distinguish
*P. dornii* from its putative close relative *P. condensata*, Lichvar noted
the basal leaves and fruits of *P. dornii* are consistently larger relative to
*P. condensata*. Additionally, though the species are geographically separated
by approximately 6 km, differences in the community composition of each taxon
were also noted. @Lichvar1983 noted *P. dornii* tended to grow with shrubs and
grasses, while *P. condensata* grew among cushion plants. *Atriplex gardneri*
(Amaranthaceae), a halophilic shrub, was also observed growing near
*P. condensata* while saline indicator species did not occur in *P. dornii*
habitats. In an area of sympatry with *P. didymocarpa* var. *integrifolia*,
Lichvar observed *P. dornii* grew in relatively looser, shaley soils as opposed
to the tighter, calcareous soils surrounding *P. didymocarpa* var.
*integrifolia*.

\paragraph{} Lastly, @Rollins1984 recognized *Physaria saximontana* Rollins as
a new species from Wyoming and Montana, distributed from southeast of the Wind
River Mountains northward through the Rocky Mountains of central Montana.
*Physaria saximontana* was distinguished from *P. didymocarpa* largely on the
basis of silique shape, where fruits of the former lack a basal sinus while the
latter tend to show a cordate, shallow basal sinus with the pedicel entering
the sinus. Additionally, the basal leaf trichomes were described as having 6-
to 8-forked primary rays with some fusion in the center. In contrast, trichomes
of the siliques show only 3- to 5-forked primary rays with no central fusion
and they spread away from the fruit. Two varieties of *P. saximontana*,
*dentata* and *saximontana*, were delineated, with *P. saximontana* var.
*dentata* Rollins described from specimens collected in Park and Glacier
Counties of Montana. The presence of broad teeth on the basal leaf margins
and a style equal to or exceeding 3 times the length of the replum was used to
separate variety *dentata* from variety *saximontana*. Additionally, the ovule
count in variety *dentata* was described as ranging from 2 to 4 per locule
while variety *saximontana* typically had 2 ovules per locule present.

\clearpage

\vspace*{-24pt}

## Geographical Distribution and Ecology

\paragraph{} *Physaria* sensu stricto occur in shrub-steppe, woodland and
montane habitats with often-arid climates and intense sunlight where they tend
to occupy siliceous soils, granitic talus or limestone shale outcrops
[@Rollins1939; @Rollins1993]. The most widespread species considered here,
*P. acutifolia* (Figure 1.1), spans Montana, Idaho, Wyoming, South Dakota,
Colorado, Utah, Nevada, Arizona and New Mexico and forms a cline with plants at
the higher elevations being smaller, often with branched, sometimes elongated,
caudices, and elongated styles (O’Kane 2010) Infraspecific taxa not recognized
by @OKane2010 include *P. acutifolia* var. *purpurea* @WelshReveal1977 of Grand
County, Utah and var. *stylosa* @Welsh1986 of northeastern Utah, the latter
previously described as *P. stylosa* by @Rollins1993.

\paragraph{} Another relatively widespread species is *P. didymocarpa*
(Hooker) A. Gray (@fig-intro-priors), which includes three
subspecies as traditionally circumscribed. The distribution of *P. didymocarpa*
subsp. *didymocarpa* extends from Alberta and British Columbia in Canada through
Washington, Idaho, Montana and Wyoming in the western U.S. [@OKane2010].
The two other subspecies include a regional endemic of Wyoming and Montana,
and a state endemic in Idaho. Known primarily from the Bighorn National Forest
of north-central Wyoming and south-central Montana, *P. didymocarpa* subsp.
*lanata* (A. Nelson) @OKane2007 has a non-overlapping distribution to the
southeast of the range of *P. didymocarpa* subsp. *didymocarpa*
[@HandleyHeidel2011]. In Idaho, *P. didymocarpa* subsp. *lyrata*
(C. L. Hitchcock) O'Kane has a narrow distribution in the Salmon River drainage
of Lemhi County [@Rollins1993; @OKane2010].

\paragraph{} Three species of *Physaria* sensu-stricto have been described from
southwestern and west-central Wyoming (@fig-intro-priors).
These include two state endemics, *P. condensata* Rollins and *P. dornii*
Lichvar, with ranges extending through the Overthrust Belt mountains of central
Uinta county and southern Lincoln county [@Fertig1998; @Fertig2002]. Both
species were formerly considered for the Endangered Species Act as C2
candidates. Currently, they are recognized as state endemics of high
conservation priority by the Wyoming National Diversity Database with
*P. dornii* given a G1/S1 heritage ranking while *P. condensata* is listed at
the G2G3/S2S3 heritage ranking (@HeidelOKane2012). *Physaria integrifolia* is a
regional endemic of western Wyoming and eastern Idaho.

{{< include /qmd/fig/_intro-priors.qmd >}}

\clearpage

\paragraph{} Three state and regional endemics have distributions through
central to eastern Wyoming (@fig-intro-priors).
Currently listed as a species of potential concern in Wyoming at the G3/S3
rank [@Heidel2014], *P. saximontana* subsp. *saximontana* is distributed in west
central Wyoming to the southeast of the Wind River Range [@Rollins1984]. The
putatively related *P. saximontana* subsp. *dentata* extends north through the
Rocky Mountains of Montana. These species occupy montane to alpine habitats
ranging from 1500 to 2700 m [@Lesica2012; @OKane2010]. *P. eburniflora* is a
species of potential concern in Wyoming given a G3/S3 heritage rank
[@Heidel2012], and a state endemic of Carbon, Fremont and Natrona Counties.
The distribution of *P. brassicoides* extends from Montana and North Dakota
south to Colorado and Nebraska at generally lower elevations, ranging from
900-1400 m [@OKane2010]. Often on bare hillsides, bluffs, riverbanks, badlands
and clay soils, its distribution is primarily non-overlapping with other
Wyoming *Physaria*.

\vspace*{18pt}

### Status of *Physaria vitulifera* in Wyoming

\paragraph{} Rollins' [-@Rollins1993] treatment of *Physaria* described
*P. vitulifera* as occurring in central Colorado. Weber and Wittman [-@Weber1996]
reported the species as East of the Front Range, on gravelly slopes of the
foothills. @Ackerfield2015 reported a similar distribution, from the South in
Fremont County to the North in Larimer County. O'Kane's [-@OKane2010] treatment
of *Physaria* follows these circumscriptions, reporting *P. vitulifera* from
only Colorado. Heidel and O'Kane [-@HeidelOKane2012] posited a need for
taxonomic work based on *Physaria* collections with annotations as
*P. vitulifera*, many of which were re-annotated by Jennings [-@Jennings2004] as
*P. acutifolia* (indicated by "\*", @tbl-intro-carbon).
In contrast to other treatments, Dorn's [-@Dorn2001] Wyoming Flora key to
*Physaria* reported the species from south-central Wyoming, distinguished by
fruits with a somewhat rectangular outline and obtuse or truncate basal apices.
It was noted the replum often attached nearly the full length of the fruit
resulting in highly reduced apices, and that basal leaf petioles tended to be
toothed, noting an occurrence on rocky slopes in or nearby mountains [@Dorn2001].
This treatment preceded the review of putative *P. vitulifera* specimens
by Jennings [-@Jennings2004]. As previously described, the taxonomic status of
*P. vitulifera* in Wyoming is unclear
[@Rollins1993; @Dorn2001; @Jennings2004; @OKane2010; @HeidelOKane2012].
A number of specimens from Carbon county in Wyoming previously annotated as
*P. vitulifera* ("*Physaria vitulifera* - Carbon")
or matching the phenotype described by Dorn [-@Dorn2001]
("*Physaria acutifolia - vitulifera*-like") are observed as
geographically separated from the Colorado Front range *P. vitulifera*
(@fig-intro-carbon).

\clearpage

{{< include /qmd/tbl/_intro-carbon.qmd >}}

```{r tbl-intro-carbon}
#| tbl-cap: Putative Carbon County, Wyoming *Physaria vitulifera* specimens.
#| tbl-subcap: >
#|   Voucher specimens within the extent ((-107.5, -105.5), (40.5, 42.0))
#|   and matched by "vitulifera" in a prior identification. Collections with a
#|   revised annotation by Jennings (2004) are distinguished by an asterisk
#|   for the lending institution of the re-annotated specimens.

# Kable of putative Wyoming P. vitulifera specimens.
carbon <-
  knitr::kable(
    x = revised,
    format = ifelse(
      test = knitr::pandoc_to() == "docx",
      yes = "pipe",
      no = knitr::pandoc_to()
    ),
    escape = FALSE,
    row.names = FALSE,
    col.names = c(
      "Most Recent Annotation",
      "Collector",
      "Number",
      "Herbarium"
    )
  )

if (knitr::pandoc_to() != "docx") {
  carbon %>%
    kableExtra::kable_styling(
      full_width = FALSE,
      latex_options= "hold_position",
      position = "center"
    ) %>%
    kableExtra::row_spec(row = 0, bold = TRUE, font_size = 10) %>%
    kableExtra::column_spec(1, border_left = TRUE) %>%
    kableExtra::column_spec(3, width = "2.5cm") %>%
    kableExtra::column_spec(4, border_right = TRUE) %>%
    kableExtra::collapse_rows(columns = 1:4)
} else {
  carbon
}
```

\vspace*{10pt}

\paragraph{} The set of sympatric specimens with a concordant phenotype show
reduced mature fruits on sigmoidal pedicels with often dentate basal leaves.
To the South, two flowering specimens (*Kastning* 1462 and *Kastning* 1725)
collected from Jackson County, Colorado show a general reduction in size with
ascending pedicels ("*Physaria*"; @fig-intro-carbon).
An additional record from Jackson County with mature fruits (*Nelson* 49286)
shows similarly ascending sigmoidal pedicels with reduced fruits. In contrast,
records further South in Middle Park of Grand County, Colorado display notably
pendent fruiting pedicels (*Nelson* 49478, *Patterson* s.n. (F-980, F-208917)),
with two having distinctly pinnatifid basal leaves
(*Patterson* s.n. (F-980, F-208917)). The *Patterson* specimens were both
collected in 1875 with location data limited to a description as from
Middle Park, in proximity to the Muddy River. Notably, North Park is separated
from Middle Park to the South by the Rabbit Ears Range which divides Grand
and Jackson Counties.

\clearpage

\vspace*{-40pt}

```{r fig-intro-carbon}
#| fig-cap: >
#|   Distribution of reported Carbon County, Wyoming *Physaria vitulifera*.
#|   SEINet specimen occurrences reveal ranges of neighboring
#|   *Physaria floribunda* subspecies southwest from Middle Park and
#|   *Physaria bellii* north of *P. vitulifera* along the eastern slope
#|   of the Front Range in Colorado. White diamonds indicate sampled DNA
#|   specimens and yellow squares denote reviewed voucher specimens of interest.
#|   SEINet specimens are otherwise indicated by smaller hollow shapes.
#| fig-scap: >
#|   *Physaria* of the Colorado Basin, Front Range, and Medicine Bow Mountains.
knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/IntroCarbonWyo",
    ext = knitr::opts_chunk$get("dev")
  )
)
```

\clearpage

\vspace*{-24pt}

## Morphology and Phylogeny

\paragraph{} Delineation of a species based on morphological characteristics
alone may result in the description of a non-monophyletic group, as it is often
not known whether the morphological characters considered are apomorphic or
plesiomorphic [@Rieseberg1994].
@Scotland2003 claim based on simulation studies that the number of
characters (e.g., sequence or morphological) needed to accurately
resolve a phylogeny exceeds by an order of magnitude those utilizable
morphological characters with unproblematic homology assessment and character
coding, noting on average relatively few characters per taxon are coded for
matrices used in morphological phylogenies.
Their suggested approach would utilize a limited number of unambiguous
morphological characters with diagnostic potential to consider monophyletic
groups in the context of molecular phylogenies.
@Wiens2004a countered that addition of morphological characters does not
necessarily increase homoplasy and can increase support and tree resolution
in *morphological* phylogeny reconstruction. Noting the lack of an apparent
null hypothesis or explicit statistical test of the scaffold approach proposed
by @Scotland2003, @Wiens2004a outlined an approach to (1) perform
separate analyses to identify incongruence between approaches; (2) complete a
combined analysis; and (3) consider strongly contested topologies as unresolved
until independent sources of evidence support a single hypothesis.

\paragraph{} For molecular phylogenies, incongruence is a notable concern for
single gene and concatenated analyses considering only a few loci, as different
genes may strongly support conflicting topologies [@Rokas2003].
While the number of possible characters available from sequence data now far
exceeds those available from morphology, conflicting genomic signal arising
from incomplete lineage sorting, horizontal gene transfer, mutation rate
heterogeneity, and choice of evolutionary model might preclude molecular data
as a sole source of evidence for phylogenetic inference [@Jenner2004].
Empirical analysis of tree-to-tree distances indicates general incongruence
between phylogenies built from only molecular or morphological data, though
either source alone provides an approximation of the underlying unknown tree
in comparison to a null hypothesis [@Pisani2007].
More recently, the precipitous drop of sequencing cost has ushered an era of
*phylogenomics* and the role of morphological phylogenetics *sensu stricto*
has been questioned as the number of number of nucleotides has far outpaced the
use of morphology in phylogenetic datasets [@LeePalci2015].
With a greater proportion of sequence data constituting simultaneous analyses
(i.e., co-estimation of phylogeny from genes and morphology; a "total evidence"
or combined approach), genomic data will increasingly influence topology such
that results approximate mapping morphology onto a molecular phylogeny (*ibid.*).
Despite this marked increase in availability of sequence data, non-phylogenetic
signal from incomplete lineage sorting remains a potential hurdle for recent
divergences of species with large effective population sizes
[@Philippe2011].

\paragraph{} Given the asymmetry between phylogenies built on molecular and
morphological data, it seems the role of morphology has increasingly been
limited to the interpretation or calibration of molecular phylogenies and
total evidence approaches with combined data [@Wanninger2015; @Giribet2015].
Morphology may also influence taxon sampling decisions and species
descriptions more broadly. Two notable taxonomic possibilities include the
recognition of infraspecific lineages as species (i.e., "splitting") or
failure to delimit closely related species as separate (i.e., "lumping";
@Faurby2016). In the case of the former, morphological variation may lead to
recognition of taxonomic entities that are not supported by a molecular phylogeny
[@Helsen2009]. For the latter, seemingly morphologically indistinguishable
species may show strong evidence of speciation by genetic evidence when
"cryptic taxa" arise from polyploid speciation [@Crowl2017].
Congruence between multiple lines of evidence may provide stronger support for
phylogenetic hypotheses and taxonomic decisions, where morphology is considered
an operational criterion for lineage separation [@deQueiroz2007]
To assess the status of *Physaria* species, observations of suspected
parsimony-informative continuous and discrete characters previously utilized in
published dichotomous keys were recorded: replum shape; ovule count; mature
fruit size, apices and trichomes; basal leaf margins; and overall habit.
Of particular interest are species previously delimited on the basis of
diagnosable monophyly from morphological characters despite the lack of
information about genetic history, cytology, or reproductive mechanisms
(e.g., *Physaria saximontana* subspecies, *Physaria eburniflora*,
*P. dornii*, *P. condensata*).

\clearpage

::: {.content-visible when-format="docx"}

{{< include /appendix/_references.qmd >}}

:::
