
\clearpage

```{r sectionNovelPhysaria}

knitr_section(
  # knitr_title = paste("Novel Wyoming", physaria(knitr_chunk)),
  knitr_title = paste("Novel Wyoming", physaria(knitr_chunk)),
  knitr_type = "section", chunk_type = knitr_chunk
)

```

\paragraph{} While the phylogeny from the sampled loci for the specimens
considered was not fully resolved, evidence for separation of the disputed
*P. vitulifera* specimens as a new taxon (henceforth *Physaria medicinae*) was
observed in morphological and ecological traits. Reviewed specimens are found
generally above 2,100m, at a higher elevation relative to nearby *P. acutifolia*
and allopatric *P. vitulifera* of Colorado 
(Figure \@ref(fig:speciesSEINet)).
Plotting the specimens from Figure \@ref(fig:FigureIntroCarbon)
over an elevation profile illustrates the geographic distance between the
putative Carbon County specimens and those of *P. vitulifera* from East of the
Front Range (Figure \@ref(fig:speciesSEINet)).
Populations of questioned specimens examined are distributed from the Sierra
Madre West of Encampment and East through the Medicine Bow Range.
After accounting for collection date, these specimens also show a marked
reduction in average mature fruit length and width relative to other nearby
species (Figure \@ref(fig:speciesPmedicinae)).

\clearpage

```{r speciesSEINet, fig.cap=caption$FigResultsSEINet, fig.scap="SEINet Distribution Map."}

# Plot of fruit dimensions by reviewed species ID and collection date.
caption$FigResultsSEINet <-
  paste(
    "Distribution of regional", physaria(knitr_chunk), "species from SEINet."
  )

knitr::include_graphics(
  path = paste0("Figs/FigResultsSEINet.", knitr_ext[[knitr_chunk]])
)

```

\clearpage

```{r speciesPmedicinae, fig.cap=caption$FigResultsPmedicinae, fig.scap="Elevation and Fruit Morphology of *Physaria medicinae*"}

caption$FigResultsPmedicinae <-
  paste("Elevation and mature fruit length differences between questioned",
        "Carbon County, WY specimens and nearby", physaria(knitr_chunk),
        "species.")

# Plot of elevation and fruit dimensions by reviewed species ID.
knitr::include_graphics(
  path = paste0("Figs/FigResultsPmedicinae.", knitr_ext[[knitr_chunk]])
)

```


```{r, eval=FALSE}
# Statistics ----

# One-way ANOVA
anovaCarbonElevation <-
  dplyr::filter(spp_carbon_wyo, !is.na(Elev_raw_max)) %>%
  aov(formula = Elev_raw_max ~ Taxon_a_posteriori, data = .) %>%
  broom::tidy()

# ANCOVA
ancovaCarbonFruits <-
  dplyr::filter(trait_carbon_fruits, !is.na(Trait_max)) %>%
  aov(data = .,
      formula = Trait_max ~ Taxon_a_posteriori + Date_md) %>% 
  broom::tidy()


```

\clearpage

```{r sectionWyomingEndemics}

knitr_section(
  knitr_title = "Southwest Wyoming Endemics",
  knitr_type = "section", chunk_type = knitr_chunk
)

```

```{r speciesSouthwestEndemics, fig.cap=caption$FigResultsSouthwestEndemics}

# Plot of *P. integrifolia*, *P. dornii*, and *P. condensata* elevation profile.
caption$FigResultsSouthwestEndemics <-
  "Distribution of southwestern Wyoming endemic *Physaria* species."

knitr::include_graphics(
  path = paste0("Figs/FigResultsSouthwestEndemics.", knitr_ext[[knitr_chunk]])
)

```

\clearpage

```{r speciesPintegrifolia, fig.cap=caption$FigResultsPintegrifolia}

# Plot of *P. integrifolia*, *P. dornii*, and *P. condensata* elevation profile.
caption$FigResultsPintegrifolia <-
  "Growth habit of southwestern Wyoming endemics."

knitr::include_graphics(
  path = paste0("Figs/FigResultsPintegrifolia.", knitr_ext[[knitr_chunk]])
)

```

\clearpage

```{r sectionPdidymocarpaComplex}

knitr_section(
  knitr_title = "*Physaria didymocarpa* Complex",
  knitr_type = "section", chunk_type = knitr_chunk
)

```

```{r speciesPdidymocarpaMap, fig.cap=caption$FigResultsPdidymocarpaMap}

# Plot of *P. didymocarpa*, *P. saximontana*, and *P. eburniflora* priors.
caption$FigResultsPdidymocarpaMap <-
  "Distribution of *P. didymocarpa* complex species."

knitr::include_graphics(
  path = paste0("Figs/FigResultsPdidymocarpaMap.", knitr_ext[[knitr_chunk]])
)

```

\clearpage

```{r speciesPdidymocarpa, fig.cap=caption$FigResultsPdidymocarpa}

# Plot of *P. didymocarpa*, *P. saximontana*, and *P. eburniflora* priors.
caption$FigResultsPdidymocarpa <-
  "Growth habit of *P. didymocarpa* complex species."

knitr::include_graphics(
  path = paste0("Figs/FigResultsPdidymocarpa.", knitr_ext[[knitr_chunk]])
)

```
