# DISCUSSION {#DISCUSSION}

\paragraph{} How species are defined is a key issue in biology, the aptly named
species problem.  Potential for amibguity in this problem was recognized early
on by @Darwin1859 [p. 46] who surmised, "No one definition has
satisfied all naturalists; yet every naturalist knows vaguely what he means when
he speaks of a species." Over the years, much consideration has been given to
this question as evidenced by the numerous species concepts described in the
literature (Reviewed by @Zachos2016). @Simpson1951 noted the importance of the
temporal dimension for the species concept debate, asserting an evolutionary 
criterion for species as, "...a phyletic lineage (ancestral-descendent sequence
of interbreeding populations) evolving independently of others, with its own
separate and unitary evolutionary role and tendencies is a basic unit of
evolution" [@Simpson1951 p. 289].  @Mayden1997 argued that while the
evolutionary species concept (ESC) described by @Simpson1951 was 
non-operational, it provided a robust theoretical framework and general 
applicability for use as a primary species concept.  With the ESC as a primary 
species concept, secondary species concepts provide mutually applicable and 
equally valid operational support for recognizing species entities
[@Mayden1997].

```{r sectionSpeciesConcept}

knitr_section(
  knitr_title = "The Species Concept",
  knitr_type = "section", chunk_type = knitr_chunk
)

```

\paragraph{} Attempting to reconcile differences among the various proposed
species concepts, @deQueiroz1998 described a general lineage concept of
species as population-level segments of evolutionary lineages to denote the
fundamental natural unit the term represents. Under this model, some concepts
describe the general lineage concept whereas other alternative species
definitions provide operational criteria by which to investigate lineage
separation. These criteria (e.g. interbreeding, monophyly, ecological niche,
cladogenesis) represent various lines of evidence for species delimitation,
where inference of an evolutionary lineage separation - that is, a species
entity - is ideally described by multiple criteria [@deQueiroz2007].
From this perspective, there is agreement about the fundamental entity
described by the term "species" despite differences in the mechanistic
hypotheses (i.e. secondary criteria) pertaining to the separation and
maintenance of the population-level evolutionary lineage segments.

```{r FigureDiscussionPosteriori, fig.cap=caption$FigureDiscussionPosteriori}

caption$FigureDiscussionPosteriori <-
  paste("Distribution of *Physaria* sensu stricto herbarium specimens",
        "by reviewed annotations.")

knitr::include_graphics(
  path = paste0("Figs/FigDiscussionPosteriori.", knitr_ext[[knitr_chunk]])
)

```

\paragraph{} Notwithstanding progress toward a universal species concept, a
complementary problem exists in species delimitation. Within the framework of
the general lineage concept, a demarcation exists between the conceptual
definition of species as evolutionary lineages and the operational criteria 
providing evidence for their existence - each of which is based on notable 
biological properties [@deQueiroz2007]. However, @Garnett2017 recently noted an
apparent lack of agreement regarding how to define lineages. The authors also
assert that international agreements, environmental legislation, and a variety
of entities concerned with biodiversity treat species as fixed entities (ibid).
Others have argued that species are far from fixed entities, and instead
represent falsifiable hypotheses with varying degrees of support
[@Thiele2002; @Hey2003; @Dayrat2005; @Raposo2017].

\paragraph{} Taxonomic nomenclature provides a necessary framework for
delineating species by discontinuous natural boundaries with utility across a 
variety of disciplines. Contending arbitrariness in the species delimitation 
process, @Garnett2017 called for global oversight and the restriction of
freedom of taxonomic action given the apparent consequences for research
programs and policy decisions dependent on accurate species estimates. However,
there may be inherent limitations in utilizing a discrete binary system to 
characterize species as entities arising from a continuous evolutionary process, 
particularly in the grey area of recent divergence following lineage splitting 
[@Zachos2018]. A degree of uncertainty is an inherent component of empirical 
scientific research, and in the case of species taxa hypotheses can result from
the dynamic demographic and evolutionary fluctuations species entities are
subject to [@Hey2003]. A recent suggestion for reducing ambiguity from species
delimitation is to explicitly cite the working species concept and include 
information concerning distribution, ecology, conservation status, and potential 
threats facing the species [@Thomson2018]. 

\paragraph{} Which species concept is applied by an investigator has
consequences for taxonomic treatments and alpha diversity [@Isaac2004]. 
Accurate species assignment and determination of species boundaries has
implications for downstream analyses. When different data support contradicting
species boundaries, a single line of evidence may fail to recover a species.
For species defined *a priori* by qualitatively combining genetic,
morphological, geographical, and ecological data, individual species assignment 
using clustering and multivariate techniques recovered species more accurately
when multiple types of data sets were included [@Edwards2014].
An "integrative taxonomy" considers naming species when supported by broad
biological evidence, where higher confidence is placed in a hypothesis supported
by different kinds of data [@Dayrat2005]. In particular, caution is warranted
for describing species on the basis of morphological diversity without support
from additional lines of evidence given the possibility of homoplasy (*ibid.*).

\paragraph{} Utilizing multiple lines of evidence (e.g genealogical,
morphological, and reproductive barriers) for delimitation may recover species
that would otherwise be missed from only a single line of evidence
[@Padial2009]. Broadly, two frameworks for an integrative taxonomy are possible: 
integration by congruence where a concordant pattern among several taxonomic 
characters indicates lineage separation, and integration by cumulation where a 
single character set may provide a good indication of lineage divergence
[@Padial2010]. The former may result in higher confidence at the cost of
decreased taxonomic resolving power, while uncritical application of the latter
may contribute to over-splitting of species taxa (ibid.). In the case of recent
radiations, integration by congruence may fail to recover species on the basis
of a strict consensus approach due to a lack of gene monophyly [@Shaffer2007].

Considerations against a generalized lineage concept (GLC)

- Hausdorf 2011
- Freudenstein 2017
- Carstens 2013 (Box 3)

