
```{r msaAppendix}

# Generate appendix sequence alignment figures
library(Biostrings)  # Pagès et al. 2017

# R 'msa' package ports LaTeX package TEXSHADE (Beitz 2000)
library(msa)  # Bodenhofer et al. 2015

# Establish a directory to store alignment appendix msa figures:
if (!dir.exists("Phys_DNA/Alignment-Figs")) {
  dir.create("Phys_DNA/Alignment-Figs")
}

# Remove old .aux and .tex files from Alignment-Figures subdirectory.
if (dir.exists("Phys_DNA/Alignment-Figs")) {
  alignment_tex_files <- list.files(path = "Phys_DNA/Alignment-Figs/", 
                                    pattern = ".aux|.tex|.fasta", 
                                    full.names = TRUE)
  invisible(file.remove(alignment_tex_files))
}

```

\clearpage

```{r msaITS}

cat("\\chapter*{APPENDIX C}
\\addcontentsline{toc}{chapter}{APPENDIX C: ITS SEQUENCE ALIGNMENT}
\\centering
ITS SEQUENCE ALIGNMENT\n",
file = "Phys_DNA/Alignment-Figs/msa_aligned_its.tex")

file_its <- paste0("Phys_DNA/Seq-Analysis/",
                   "3.unwrapped_fastas_single/rITS-unwrapped_single.fasta")

aligned_its <- readDNAMultipleAlignment(filepath = file_its, format = "fasta")

msaPrettyPrint(aligned_its, y = c(1, 135), output = "tex",
               alFile = "Phys_DNA/Alignment-Figs/aligned_its.fasta",
               file = "Phys_DNA/Alignment-Figs/aligned_its.tex",
               showLogo = "none", askForOverwrite = FALSE, showLegend = FALSE,
               furtherCode = "\\showruler{1}{top}")

system("./Phys_DNA/msa_format.sh \\
       -F Phys_DNA/Alignment-Figs/aligned_its.tex -m 7in -r ROTATE >> \\
       Phys_DNA/Alignment-Figs/msa_aligned_its.tex")

```

\include{Phys_DNA/Alignment-Figs/msa_aligned_its}

\begin{landscape}

```{r msaITSLandscape}

msaPrettyPrint(aligned_its, y = c(136, max(aligned_its@unmasked@ranges@width)),
               output = "tex",
               alFile = "Phys_DNA/Alignment-Figs/aligned_its.fasta",
               file = "Phys_DNA/Alignment-Figs/aligned_its_end.tex",
               showLogo = "none", askForOverwrite = FALSE, showLegend = FALSE,
               furtherCode = "\\showruler{1}{top}")

system("./Phys_DNA/msa_format.sh \\
       -F Phys_DNA/Alignment-Figs/aligned_its_end.tex -r LANDSCAPE > \\
       Phys_DNA/Alignment-Figs/msa_aligned_its_end.tex")

```

\include{Phys_DNA/Alignment-Figs/msa_aligned_its_end}

\end{landscape}

```{r msaRPS}

cat("\\chapter*{APPENDIX D}
\\addcontentsline{toc}{chapter}{APPENDIX D: \\textit{rps} SEQUENCE ALIGNMENT}
\\centering
\\textit{rps} SEQUENCE ALIGNMENT\n",
file = "Phys_DNA/Alignment-Figs/msa_aligned_rps.tex")

file_rps <- paste0("Phys_DNA/Seq-Analysis/",
                   "3.unwrapped_fastas_single/rps-unwrapped_single.fasta")

aligned_rps <- readDNAMultipleAlignment(filepath = file_rps, format = "fasta")

msaPrettyPrint(aligned_rps, y = c(1, 135), output = "tex",
               alFile = "Phys_DNA/Alignment-Figs/aligned_rps.fasta",
               file = "Phys_DNA/Alignment-Figs/aligned_rps.tex",
               showLogo = "none", askForOverwrite = FALSE, showLegend = FALSE,
               furtherCode = "\\showruler{1}{top}")

system("./Phys_DNA/msa_format.sh \\
       -F Phys_DNA/Alignment-Figs/aligned_rps.tex -m 7in -r ROTATE >> \\
       Phys_DNA/Alignment-Figs/msa_aligned_rps.tex")

```

\include{Phys_DNA/Alignment-Figs/msa_aligned_rps}

\begin{landscape}

```{r msaRPSLandscape}

msaPrettyPrint(aligned_rps, y = c(136, max(aligned_rps@unmasked@ranges@width)),
               output = "tex",
               alFile = "Phys_DNA/Alignment-Figs/aligned_rps.fasta",
               file = "Phys_DNA/Alignment-Figs/aligned_rps_end.tex",
               showLogo = "none", askForOverwrite = FALSE, showLegend = FALSE,
               furtherCode = "\\showruler{1}{top}")

system("./Phys_DNA/msa_format.sh \\
       -F Phys_DNA/Alignment-Figs/aligned_rps_end.tex -r LANDSCAPE > \\
       Phys_DNA/Alignment-Figs/msa_aligned_rps_end.tex")

```

\include{Phys_DNA/Alignment-Figs/msa_aligned_rps_end}

\end{landscape}

```{r msaYCF1}

cat("\\chapter*{APPENDIX E}
\\addcontentsline{toc}{chapter}{APPENDIX E: \\textit{ycf1} SEQUENCE ALIGNMENT}
\\centering
\\textit{ycf1} SEQUENCE ALIGNMENT\n",
file = "Phys_DNA/Alignment-Figs/msa_aligned_ycf1.tex")

file_ycf1 <- paste0("Phys_DNA/Seq-Analysis/",
                   "3.unwrapped_fastas_single/ycf1-unwrapped_single.fasta")

aligned_ycf1 <- readDNAMultipleAlignment(filepath = file_ycf1, format = "fasta")

msaPrettyPrint(aligned_ycf1, y = c(1, 135), output = "tex",
               alFile = "Phys_DNA/Alignment-Figs/aligned_ycf1.fasta",
               file = "Phys_DNA/Alignment-Figs/aligned_ycf1.tex",
               showLogo = "none", askForOverwrite = FALSE, showLegend = FALSE,
               furtherCode = "\\showruler{1}{top}")

system("./Phys_DNA/msa_format.sh \\
       -F Phys_DNA/Alignment-Figs/aligned_ycf1.tex -m 7in -r ROTATE >> \\
       Phys_DNA/Alignment-Figs/msa_aligned_ycf1.tex")

```

\include{Phys_DNA/Alignment-Figs/msa_aligned_ycf1}

\begin{landscape}

```{r msaYCF1Landscape}

msaPrettyPrint(aligned_ycf1, y = c(136, max(aligned_ycf1@unmasked@ranges@width)),
               output = "tex",
               alFile = "Phys_DNA/Alignment-Figs/aligned_ycf1.fasta",
               file = "Phys_DNA/Alignment-Figs/aligned_ycf1_end.tex",
               showLogo = "none", askForOverwrite = FALSE, showLegend = FALSE,
               furtherCode = "\\showruler{1}{top}")

system("./Phys_DNA/msa_format.sh \\
       -F Phys_DNA/Alignment-Figs/aligned_ycf1_end.tex -r LANDSCAPE > \\
       Phys_DNA/Alignment-Figs/msa_aligned_ycf1_end.tex")

```

\include{Phys_DNA/Alignment-Figs/msa_aligned_ycf1_end}

\end{landscape}

