
```{r, eval=FALSE}

# One-way ANOVA
anovaPhysariaElevation <- herbarium_specimens %>%
  filter(!is.na(Taxon_a_posteriori) & !is.na(Elev_raw_max)) %>%
  dplyr::filter(., !is.na(Elev_raw_max)) %>%
  aov(formula = Elev_raw_max ~ Taxon_a_posteriori, data = .) %>%
  broom::tidy()

# ANCOVA
ancovaPhysariaFruits <- herbarium_specimens %>%
  dplyr::filter(!is.na(Ovule_number), !is.na(Taxon_a_posteriori)) %>%
  bind_cols(range_split(trait_tbl = ., split_var = "Ovule_number")) %>%
  aov(data = .,
      formula = Ovule_number_max ~ Taxon_a_posteriori + Date_md) %>%
  broom::tidy()

anovaPhysariaStems <- herbarium_specimens %>%
  dplyr::filter(!is.na(Stem_length_dm), !is.na(Taxon_a_posteriori)) %>%
  bind_cols(range_split(trait_tbl = ., split_var = "Stem_length_dm")) %>%
  aov(data = .,
      formula = Stem_length_dm_max ~ Taxon_a_posteriori) %>%
  broom::tidy()

```

\clearpage

\vspace*{-18pt}

```{r sectionTraitPlots}

Thesis::knitr_section(
  knitr_title = "Trait Distributions",
  knitr_type = "section", chunk_type = knit_variables$pandoc
)

```

\paragraph{} Records of trait observations from specimen vouchers were used for
exploratory data analysis of species hypotheses based on morphological and
ecological data. Discrete observations of ovules (or *funiculi*: persistent,
stalk-like protrusions along the replum) generally follow one of two trends:
taxa with predominantly 2 ovules per locules, and those with commonly 3-4 or
more ovules *not restricted* to the apex of the replum
(Panel A, Figure \@ref(fig:FigResultsPhysariaOvules)).
*Physaria acutifolia*, *P. brassicoides*, *P. didymocarpa* subsp.
*lanata*, *P. vitulifera*, and individuals of the Sierra Madre and Medicine Bow
mountains comprise the former group. The latter group includes *P. didymocarpa*
subspecies, previously-annotated *P. saximontana* subspecies
(Panel B, Figure \@ref(fig:FigResultsPhysariaOvules)),
*P. eburniflora*, and the southwest Wyoming endemics *P. integrifolia*,
*P. dornii*, and *P. condensata*.
Fruit length from `r physaria(knitr_chunk = knit_variables$pandoc)`
individuals of the Sierra Madre and Medicine Bow mountain ranges are
highly reduced relative to other species
(*P.* 'medicinae' in Panel A, Figure \@ref(fig:FigResultsPhysariaFruits)).
When accounting for collection date, these individuals form fruits through
June similar to nearby *P. acutifolia*
(Panel B, Figure \@ref(fig:FigResultsPhysariaFruits)). In contrast,
observations indicate individuals of *P. integrifolia* and *P. didymocarpa*
subsp. *lyrata* may form slightly larger fruits later into the season.
Other subspecies of *Physaria didymocarpa*, subsp. *lanata* and subsp.
*didymocarpa*, fruit into July and August with sizes in the range of 7-12 mm.

\vspace*{24pt}

```{r FigResultsPhysariaOvules, fig.cap=caption$FigResultsPhysariaOvules, fig.scap="Maximum ovule counts per locule"}

caption$FigResultsPhysariaOvules <-
  "Maximum ovule counts per locule."

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/ResultsPhysariaOvules",
    ext = knit_variables$graphics
  )
)

```

\clearpage
\vfill

```{r FigResultsPhysariaFruits, fig.cap=caption$FigResultsPhysariaFruits, fig.scap="Fruit size observations from reviewed specimen vouchers"}

caption$FigResultsPhysariaFruits <-
  "Observations of fruit size from reviewed specimen vouchers."

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/ResultsPhysariaFruits",
    ext = knit_variables$graphics
  )
)

```

\vfill
\clearpage

```{r statisticsPhysariaElevation}

results <- list()

results$elevation$specimens <- Thesis::herbarium_specimens %>%
  Thesis::filter_reviewed(specimen_tbl = .) %>%
  dplyr::filter(
    !is.na(Elev_raw_max),
    !is.na(Taxon_a_posteriori),
    !grepl("^Physaria$", .data$Taxon_a_posteriori)
  ) %>%
  dplyr::mutate(
    elevation = Elev_raw_max,
    taxon = as.factor(Taxon_a_posteriori)
  )

results$elevation$aov <- results$elevation$specimens %>%
   aov(elevation ~ taxon, data = .)

results$elevation$shapiro <- shapiro.test(results$elevation$aov$residuals) %>%
  broom::tidy() %>%
  dplyr::pull("p.value")
# Patrick Royston (1982). An extension of Shapiro and Wilk's W test for normality to large samples. Applied Statistics, 31, 115–124. doi: 10.2307/2347973.
# Patrick Royston (1982). Algorithm AS 181: The W test for Normality. Applied Statistics, 31, 176–180. doi: 10.2307/2347986.
# Patrick Royston (1995). Remark AS R94: A remark on Algorithm AS 181: The W test for normality. Applied Statistics, 44, 547–551. doi: 10.2307/2986146.

results$elevation$levene <- results$elevation$specimens %>%
  car::leveneTest(elevation ~ taxon, data = .) %>%
  broom::tidy() %>%
  dplyr::pull("p.value")
# Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
# Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.

results$elevation$kruskal <- results$elevation$specimens %>%
  kruskal.test(elevation ~ taxon, data = .) %>%
  broom::tidy() %>%
  dplyr::pull("p.value")

```

\paragraph{} Maximum values were obtained from specimen voucher elevation
ranges (Figure \@ref(fig:FigResultsPhysariaElevation)). In total, elevation
records from `r nrow(results$elevation$specimens)` specimens were included.
Elevation data were non-normal according to the Shapiro-Wilk normality test 
(p=`r results$elevation$shapiro`), with each of *P. didymocarpa* subspecies
*lanata* and *lyrata* showing bimodal distributions
(Panels A and B, Figure \@ref(fig:FigResultsPhysariaElevation)).
Additionally, variances were unequal according to Levene's test of variance
homogeneity (p=`r results$elevation$levene`). Non-parametric Kruskal-Wallis 
rank sum test indicates the elevation medians are unequal
(p=`r results$elevation$kruskal`). These ranges reveal a relatively narrow 
distribution for Wyoming endemic species *P. condensata*, *P. dornii*, and
*P. eburniflora* (Panels B and C, Figure \@ref(fig:FigResultsPhysariaElevation)).
Individuals from the Sierra Madre and Medicine Bow ranges had the highest median
distribution at greater than 8,000 ft.

```{r FigResultsPhysariaElevation, fig.cap=caption$FigResultsPhysariaElevation, fig.scap="Maximum elevation records from specimen vouchers"}

caption$FigResultsPhysariaElevation <-
  paste(
    "Elevation records from specimen vouchers.",
    "Panel A: Violin plots with jittered points by reviewed annotation.",
    "Panel B: Ridgeline plot with relative maximum elevation distributions.",
    "Panel C: Summary statistics for median of maximum elevation (ft.),",
    "median absolute deviation (MAD), and total specimens (n)."
  )

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/ResultsPhysariaElevation",
    ext = knit_variables$graphics
  )
)

```

\clearpage

\paragraph{} Measurements of basal leaf length, stem length, and the degree to
which the inflorescence exceeds the basal rosette reveal three species with
condensed growth (Figure \@ref(fig:FigResultsPhysariaHabit)). Both of the
southwest Wyoming endemics *P. dornii* and *P. condensata* have relatively
short stem lengths with inflorescences barely exceeding the basal rosette
(Panels B and C Figure \@ref(fig:FigResultsPhysariaHabit)), where the former
is distinguished by larger basal leaves
(Panels A Figure \@ref(fig:FigResultsPhysariaHabit)). The other condensed growth
habit species from central Wyoming *P. eburniflora* is similarly distinguished
from other `r physaria(knitr_chunk = knit_variables$pandoc)` species by
inflorescences barely exceeding basal rosettes and relatively shorter stems.

\vspace*{24pt}

```{r FigResultsPhysariaHabit, fig.cap=caption$FigResultsPhysariaHabit, fig.scap="Growth habit of \\textit{Physaria} species"}

caption$FigResultsPhysariaHabit <-
  physaria(knitr_chunk = knit_variables$pandoc) %>%
  paste("Growth habit of", ., "species.")

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/ResultsPhysariaHabit",
    ext = knit_variables$graphics
  )
)

```
