---
bibliography: "references.bib"
csl: council-of-science-editors-author-date.csl
---

```{r, include=FALSE}
WORD_DOC <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx"
```

```{r, child=if (WORD_DOC) '_setup.Rmd'}
```

# DISCUSSION {#DISCUSSION}

\paragraph{} How species are defined is a key issue in biology, the aptly named
species problem.  Potential for amibguity in this problem was recognized early
on by @Darwin1859 [p. 46] who surmised, "No one definition has
satisfied all naturalists; yet every naturalist knows vaguely what he means when
he speaks of a species." Over the years, much consideration has been given to
this question as evidenced by the numerous species concepts described in the
literature (Reviewed by @Zachos2016). @Simpson1951 noted the importance of the
temporal dimension for the species concept debate, asserting an evolutionary
criterion for species as, "...a phyletic lineage (ancestral-descendent sequence
of interbreeding populations) evolving independently of others, with its own
separate and unitary evolutionary role and tendencies is a basic unit of
evolution" [@Simpson1951 p. 289].  @Mayden1997 argued that while the
evolutionary species concept (ESC) described by @Simpson1951 was
non-operational, it provided a robust theoretical framework and general
applicability for use as a primary species concept.  With the ESC as a primary
species concept, secondary species concepts provide mutually applicable and
equally valid operational support for recognizing species entities
[@Mayden1997].

```{r sectionSpeciesConcept}

knitr_section(
  knitr_title = "The Species Concept",
  knitr_type = "section", chunk_type = knit_variables$pandoc
)

```

\paragraph{} Attempting to reconcile differences among the various proposed
species concepts, @deQueiroz1998 described a general lineage concept of
species as population-level segments of evolutionary lineages to denote the
fundamental natural unit the term represents. Under this model, some concepts
describe the general lineage concept whereas other alternative species
definitions provide operational criteria by which to investigate lineage
separation. These criteria (e.g. interbreeding, monophyly, ecological niche,
cladogenesis) represent various lines of evidence for species delimitation,
where inference of an evolutionary lineage separation - that is, a species
entity - is ideally described by multiple criteria [@deQueiroz2007].
From this perspective, there is agreement about the fundamental entity
described by the term "species" despite differences in the mechanistic
hypotheses (i.e. secondary criteria) pertaining to the separation and
maintenance of the population-level evolutionary lineage segments.

```{r FigureDiscussionPosteriori, fig.cap=caption$FigureDiscussionPosteriori}

caption$FigureDiscussionPosteriori <-
  paste("Distribution of *Physaria* sensu stricto herbarium specimens",
        "by reviewed annotations.")

knitr::include_graphics(
  path = system.file(
    fs::path("figures/DiscussionPosteriori",
             ext = knit_variables$graphics),
    package = "Thesis"
  )
)

```

\paragraph{} Notwithstanding progress toward a universal species concept, a
complementary problem exists in species delimitation. Within the framework of
the general lineage concept, a demarcation exists between the conceptual
definition of species as evolutionary lineages and the operational criteria
providing evidence for their existence - each of which is based on notable
biological properties [@deQueiroz2007]. However, @Garnett2017 recently noted an
apparent lack of agreement regarding how to define lineages. The authors also
assert that international agreements, environmental legislation, and a variety
of entities concerned with biodiversity treat species as fixed entities (ibid).
Others have argued that species are far from fixed entities, and instead
represent falsifiable hypotheses with varying degrees of support
[@Thiele2002; @Hey2003; @Dayrat2005; @Raposo2017].

\paragraph{} Taxonomic nomenclature provides a necessary framework for
delineating species by discontinuous natural boundaries with utility across a
variety of disciplines. Contending arbitrariness in the species delimitation
process, @Garnett2017 called for global oversight and the restriction of
freedom of taxonomic action given the apparent consequences for research
programs and policy decisions dependent on accurate species estimates. However,
there may be inherent limitations in utilizing a discrete binary system to
characterize species as entities arising from a continuous evolutionary process,
particularly in the grey area of recent divergence following lineage splitting
[@Zachos2018]. A degree of uncertainty is an inherent component of empirical
scientific research, and in the case of species taxa hypotheses can result from
the dynamic demographic and evolutionary fluctuations species entities are
subject to [@Hey2003]. A recent suggestion for reducing ambiguity from species
delimitation is to explicitly cite the working species concept and include
information concerning distribution, ecology, conservation status, and potential
threats facing the species [@Thomson2018].

\paragraph{} Which species concept is applied by an investigator has
consequences for taxonomic treatments and alpha diversity [@Isaac2004].
Accurate species assignment and determination of species boundaries has
implications for downstream analyses. When different data support contradicting
species boundaries, a single line of evidence may fail to recover a species.
For species defined *a priori* by qualitatively combining genetic,
morphological, geographical, and ecological data, individual species assignment
using clustering and multivariate techniques recovered species more accurately
when multiple types of data sets were included [@Edwards2014].
An "integrative taxonomy" considers naming species when supported by broad
biological evidence, where higher confidence is placed in a hypothesis supported
by different kinds of data [@Dayrat2005]. In particular, caution is warranted
for describing species on the basis of morphological diversity without support
from additional lines of evidence given the possibility of homoplasy (*ibid.*).

\paragraph{} Utilizing multiple lines of evidence (e.g genealogical,
morphological, and reproductive barriers) for delimitation may recover species
that would otherwise be missed from only a single line of evidence
[@Padial2009]. Broadly, two frameworks for an integrative taxonomy are possible:
integration by congruence where a concordant pattern among several taxonomic
characters indicates lineage separation, and integration by cumulation where a
single character set may provide a good indication of lineage divergence
[@Padial2010]. The former may result in higher confidence at the cost of
decreased taxonomic resolving power, while uncritical application of the latter
may contribute to over-splitting of species taxa (ibid.). In the case of recent
radiations, integration by congruence may fail to recover species on the basis
of a strict consensus approach due to a lack of gene monophyly [@Shaffer2007].

Considerations against a generalized lineage concept (GLC)

- Hausdorf 2011
- Freudenstein 2017
- Carstens 2013 (Box 3)

\clearpage

\vspace*{-24pt}

```{r sectionDichotomousKey}

knitr_section(
  knitr_title = "Dichotomous Key",
  knitr_type = "section", chunk_type = knit_variables$pandoc
)

```

\singlespacing

\begin{enumerate}

  \item[1. ]
  Fruits didymous; replum constricted, linear; ovules 2 per locule,
  restricted to replum apex.

    \begin{enumerate}

  \item[2. ] Sinuses of mature fruits equal; from northern Wind River and
  Green River Basins in Wyoming south through the Colorado River Basin
  \dotfill \textit{Physaria acutifolia} Rydberg

  \item[2. ] Sinuses of mature fruits unequal in size.

        \begin{enumerate}

  \item[3. ] Fruits cordate; eastern Wyoming north to Montana and western
  North and South Dakota \dotfill \textit{Physaria brassicoides} Rydberg

  \item[3. ] Fruits angular, often wrinkled, with truncate basal sinuses.

            \begin{enumerate}

  \item[4. ] Mature fruits > 8 mm; pedicels recurved, basal leaves
  pinnatifid; north from Middle Park Colorado \dotfill
  \textit{Physaria floribunda} Rydberg

  \item[4. ] Mature fruits < 8 mm; pedicels ascending or sigmoidal.

                \begin{enumerate}

  \item[5. ] Mature fruits 4-8 mm; basal leaves pandurate, often with
  dentate margins; Eastern Front Range, Colorado
  \dotfill \textit{Physaria vitulifera} Rydberg

  \item[5. ] Mature fruits < 5 mm; plants alpine; Medicine Bow Mountains
  and Sierra Madre Range of Wyoming, NC Colorado
  \dotfill \textit{Physaria medicinaea} Ratcliff

                \end{enumerate}

            \end{enumerate}

        \end{enumerate}

    \end{enumerate}

\end{enumerate}

\begin{enumerate}

  \item[1. ] Fruits didymous; replum wider, oblong or elliptic to oblanceolate
  or obovate; ovules (2-)4(-6) per locule, not restricted to apex of replum.

    \begin{enumerate}

  \item[6. ] Racemes equal to or barely exceeding the rosette.

        \begin{enumerate}

  \item[7. ] Petals whitish, pale yellow; basal sinuses of mature fruits
  shallow or absent with spreading trichomes; Wyoming state endemic of
  Fremont, Natrona and Carbon counties
  \dotfill \textit{Physaria eburniflora} Rollins

  \item[7. ] Petals yellow; basal sinuses of mature fruits equally incised
  above and below with appressed trichomes.

            \begin{enumerate}

  \item[8. ] Basal leaves $\geq$ 2cm; mature fruits generally $\geq$ 1 x 1 cm;
  Wyoming state endemic of Lincoln and Uinta counties
  \dotfill \textit{Physaria dornii} Lichvar

  \item[8. ] Basal leaves more often $\leq$ 2cm; mature fruits $\leq$ 1 x 1 cm;
  Wyoming state endemic of Sublette, Lincoln and Uinta counties
  \dotfill \textit{Physaria condensata} Rollins

            \end{enumerate}

        \end{enumerate}

  \item[6. ] Racemes exceeding the basal rosette.

        \begin{enumerate}

  \item[9. ] Basal sinuses of fruits absent or less often shallow, apical
  sinuses deeper; > 2 ovules per locule   %%% \newline \hspace*{0pt} %%%
  \dotfill \textit{Physaria didymocarpa} (Hooker) A. Gray
  subsp. \textit{didymocarpa}

  \item[9. ]  Fruit sinuses generally equal, less often basally shallow;

            \begin{enumerate}

  \item[10. ] Basal leaf margins generally dentate or toothed;
  mature fruit trichomes spreading; 2 ovules per locule;
  regional endemic of the Big Horn Mountains in Wyoming and Montana
  \dotfill \textit{Physaria didymocarpa} (Hooker) A. Gray
  subsp. \textit{lanata} (A. Nelson) O'Kane

  \item[10. ] Basal leaf margins more often entire, occasionally toothed;
  mature fruit trichomes appressed; > 2 ovules per locule;
  regional endemic of SW Wyoming and eastern Idaho
  \dotfill \textit{Physaria integrifolia} (Rollins) Lichvar

            \end{enumerate}

        \end{enumerate}

    \end{enumerate}

\end{enumerate}

\doublespacing

\clearpage

\vspace*{-24pt}

```{r sectionDescriptions}

knitr_section(
  knitr_title = "Species Descriptions",
  knitr_type = "section", chunk_type = knit_variables$pandoc
)

```

\vspace*{-18pt}

```{r subsectionDescriptionPacutifolia}

knitr_section(
  knitr_title = "*Physaria acutifolia* Rydberg",
  knitr_type = "subsection", chunk_type = knit_variables$pandoc
)

```

\paragraph{} **Perennials**; caudex simple, less often branched, (tufted or condensed);
 densely pubescent, with appressed stellate trichomes, rays bifurcating, umbonate.
**Stems** several, 0.4-1.2(-2) dm, exceeding a basal rosette,
 decumbent to ascending.
**Basal leaves**: rhombic-lenticular, less often suborbicular or obovate,
 2-5(-9) cm, margins entire, occasionally toothed, (apex acute),
 petioles slender, tapering from basal leaf.
**Cauline leaves**: oblanceolate, (spatulate), 1-3(-4) cm,
 margins entire, apex acute.
**Racemes** elongated, less often congested to subumbellate.
**Fruiting pedicels** erect or ascending, often sigmoidal, 6-12 mm.
**Flowers**: petals yellow, spatulate, 7-10(-12) mm;
 sepals lanceolate, 5-7(-8) mm.
**Fruits** didymous, inflated, reniform to suborbicular, 5-10(-15) x 6-12(-20) mm,
 apices more or less equal; valves pubescent with trichomes appressed;
 replum linear, constricted, (oblong), rarely oblong or slightly oblanceolate;
 ovules 2(-4) per locule, restricted to apex of the replum; style 4-6(-8) mm.
**Seeds** brown, flattened; 2-3(-4) mm.

```{r subsectionDescriptionPbrassicoides}

knitr_section(
  knitr_title = "*Physaria brassicoides* Rydberg",
  knitr_type = "subsection", chunk_type = knit_variables$pandoc
)

```

\paragraph{} **Perennials**; caudex simple, less often branched;
 densely pubescent, with appressed stellate trichomes, rays bifurcating, umbonate.
**Stems** several, 0.6-1.3(-3) dm, exceeding a basal rosette,
 decumbent to ascending or erect.
**Basal leaves**: rhombic to obovate, 3-7(-10) cm,
 margins often dentate, otherwise entire,
 petioles slender, tapering from basal leaf.
**Cauline leaves**: oblanceolate to spatulate, 1-3(-5) cm,
 margins entire, apex acute.
**Racemes** elongated, less often congested.
**Fruiting pedicels** ascending, generally sigmoidal, 5-12 mm.
**Flowers**: petals yellow, spatulate, 7-9(-11) mm;
 sepals lanceolate, 5-7 mm.
**Fruits** didymous, distinctly cordate, 5-9(-13) x 7-14(-22) mm,
 lacking a basal sinus with a broad apical sinus;
 valves pubescent with trichomes spreading, generally furcate;
 replum linear, often with an obtuse apex;
 ovules 2 per locule, restricted to apex of the replum; style 4-6(-8) mm.
**Seeds** brown, flattened; 2-3(-4) mm.

```{r subsectionDescriptionPcondensata}

knitr_section(
  knitr_title = "*Physaria condensata* Rollins",
  knitr_type = "subsection", chunk_type = knit_variables$pandoc
)

```

\paragraph{} **Perennials**; caudex simple, (branched); plants compact, rosulate;
 densely pubescent, with appressed stellate trichomes, rays bifurcating, (umbonate).
**Stems** several, 0.1-0.4(-0.6) dm, often equal to basal rosette,
 decumbent to ascending.
**Basal leaves**: rhombic-lenticular to obovate, 0.5-2.5 cm,
 margins entire, tapering from basal leaf.
**Cauline leaves**: oblanceolate, 0.5-1.0(-1.5) cm, margins entire.
**Racemes** subumbellate to congested.
**Fruiting pedicels** erect to ascending, 5-10 mm.
**Flowers**: petals yellow, spatulate, 6-8(-9) mm;
 sepals lanceolate, 4-5(-6) mm.
**Fruits** didymous, 4-7(-10) x 4-8(-12) mm, sinuses equal above and below;
 valves pubescent with trichomes appressed, bifurcate, (umbonate);
 replum obovate to oblanceolate, apex obtuse;
 ovules 3-4(-5) per locule; style 3-5(-6) mm.
**Seeds** brown, flattened; 2-3 mm.

```{r subsectionDescriptionPdidymocarpaSspDidymocarpa}

knitr_section(
  knitr_title = "*Physaria didymocarpa* (Hooker) A. Gray subsp. *didymocarpa*",
  knitr_type = "subsection", chunk_type = knit_variables$pandoc
)

```

\paragraph{} **Perennials**; caudex simple, less often branched;
 densely pubescent, with appressed stellate trichomes, rays furcate or bifurcate.
**Stems** several to numerous, (few), 0.3-1.0(-2) dm, exceeding a basal rosette,
 decumbent to ascending.
**Basal leaves**: obovate, rhombic to suborbicular, 1.5-4(-8) cm,
 margins more often dentate than entire, (apex acute),
 petioles slender, tapering from blade, (abruptly narrowing), (lyrate).
**Cauline leaves**: oblanceolate, (spatulate), 1-2(-3) cm,
 margins entire, (dentate).
**Racemes** elongated or congested to subumbellate.
**Fruiting pedicels** erect or ascending, curved to sigmoidal, 8-12 mm.
**Flowers**: petals yellow, spatulate, 7-11(-15) mm;
 sepals lanceolate, 5-7(-10) mm.
**Fruits** didymous, inflated, reniform to suborbicular,
 5-10(-20) x 5-15(-25) mm, basal sinus shallow or absent, apical sinus deep;
 valves pubescent with spreading trichomes, rays simple or branched, furcate;
 replum elliptic (oblong) to oblanceolate or obovate;
 ovules (2-)4(-6) per locule; style 5-7(-10) mm.
**Seeds** brown, flattened; 2-3(-4) mm.

```{r subsectionDescriptionPdornii}

knitr_section(
  knitr_title = "*Physaria dornii* Lichvar",
  knitr_type = "subsection", chunk_type = knit_variables$pandoc
)

```

\paragraph{} **Perennials**; caudex simple; plants compact, rosulate;
 densely pubescent, with appressed stellate trichomes, rays bifurcating, (umbonate).
**Stems** several, 0.2-0.6(-1.0) dm, barely exceeding a basal rosette, erect.
**Basal leaves**: rhombic-lenticular to suborbicular, 2-5(-7.5) cm,
 margins entire, tapering from basal leaf.
**Cauline leaves**: oblanceolate, 0.8-2.0(-2.5) cm, margins entire.
**Racemes** subumbellate to congested.
**Fruiting pedicels** erect to ascending, 7-18 mm.
**Flowers**: petals yellow, spatulate, 7-9(-10) mm;
 sepals lanceolate, 5-6 mm.
**Fruits** didymous, 6-12(-16) x 8-16(-18) mm, sinuses equal above and below;
 valves pubescent with trichomes appressed, more often furcate, (umbonate);
 replum oblanceolate to obovate or elliptic; ovules 2-4(-6) per locule;
 style 3-5(-7) mm.
**Seeds** brown, flattened; 2-3 mm.


```{r subsectionDescriptionPintegrifolia}

knitr_section(
  knitr_title = "*Physaria integrifolia* (Rollins) Lichvar",
  knitr_type = "subsection", chunk_type = knit_variables$pandoc
)

```

\paragraph{} **Perennials**; caudex simple or often branched;
 densely pubescent, with appressed stellate trichomes, rays bifurcating, (umbonate).
**Stems** several, 0.5-1.5(-2) dm, exceeding the basal rosette,
 decumbent to ascending.
**Basal leaves**: rhombic-orbicular to obovate, 2-5(-10) cm,
 margins entire, (toothed), petiole slender, tapering from blade.
**Cauline leaves**: oblanceolate, 1.0-2.5(-5) cm, margins entire.
**Racemes** elongated, less often congested or subumbellate.
**Fruiting pedicels** erect to ascending, less often sigmoidal, 7-11 mm.
**Flowers**: petals yellow, spatulate, 8-11(-13) mm;
 sepals lanceolate, 5-8(-10) mm.
**Fruits** didymous, inflated, 6-13(-20) x 8-17(-30) mm,
 sinuses equal above and below, (shallow basal sinus);
 valves pubescent with trichomes appressed, furcate or bifurcate, (umbonate);
 replum oblanceolate to obovate, elliptic or oblong; ovules 2-4(-6) per locule;
 style 5-7(-9) mm.
**Seeds** brown, flattened; 2-3 mm.

```{r subsectionDescriptionPmedicinae}

knitr_section(
  knitr_title = "*Physaria medicinae* Ratcliff, Sp. Nov.",
  knitr_type = "subsection", chunk_type = knit_variables$pandoc
)

```

\paragraph{} **Perennials**; caudex simple, (branched);
 densely pubescent, with appressed stellate trichomes, rays bifurcating, (umbonate).
**Stems** several, 0.3-1(-1.5) dm, prostrate to decumbent, less often ascending.
**Basal leaves**: rhombic to suborbicular, obovate or ovate; 2-5(-6) cm,
 margins entire, less often dentate; petiole slender, abruptly narrowing from blade.
**Cauline leaves**: oblanceolate to spatulate, 0.8-2(-2.5) cm, margins entire.
**Racemes** elongated, (subumbellate).
**Fruiting pedicels** erect to ascending, sigmoidal.
**Flowers**: petals yellow, spatulate, 6-9(-11) mm;
 sepals lanceolate, 4-6(-7) mm.
**Fruits** didymous, angular, often truncate; 3-4(-6) mm x 3-6(-7) mm,
 sinuses more or less equal; valves pubescent with trichomes appressed,
 rays furcate to less often bifurcate; replum linear, (constricted),
 (with simple trichomes); ovules 2 per locule; style 3-6(-7) mm.
**Seeds** brown, flattened, 2(-3) mm.


```{r subsectionDescriptionPvitulifera}

knitr_section(
  knitr_title = "*Physaria vitulifera* Rydberg",
  knitr_type = "subsection", chunk_type = knit_variables$pandoc
)

```

\paragraph{} **Perennials**; caudex simple; densely pubescent,
 with appressed stellate trichomes, rays bifurcating.
**Stems** several, 0.5-1.7(-2.5) dm, exceeding the basal rosette,
 decumbent to ascending.
**Basal leaves**: often pandurate, rhombic or elliptic, 2-6(-10) cm,
 margins dentate, (entire), petiole slender, tapering from blade,
 (abruptly narrowing), often lyrate.
**Cauline leaves**: oblanceolate to spatulate, 1-2.5(-3.5) cm, margins entire.
**Racemes** generally elongated, less often subumbellate.
**Fruiting pedicels** sigmoidal or otherwise ascending, 6–10 mm.
**Flowers**: petals yellow, spatulate, 6-9(-11) mm;
 sepals lanceolate, 4-6(-8) mm.
**Fruits** didymous, angular, (wrinkled); 4-6(-8) x 5-10(-15) mm,
 basal sinus truncate, apical deeper;
 valves pubescent with trichomes appressed to spreading, furcate, (bifurcate);
 replum linear; ovules 2 per locule; style 4-6(-8) mm.
**Seeds** brown, flattened; 2-3(-4) mm.

```{r, child=if (WORD_DOC) '05-APPENDICES.Rmd'}
```

