
```{r SectionNovelPhysaria}

Thesis::knitr_section(
  knitr_title = paste(
    "Southcentral Wyoming",
    physaria(knit_variables$pandoc), "Lineage"
  ),
  knitr_type = "section", chunk_type = knit_variables$pandoc
)

```

\paragraph{} Divergence of a novel *Physaria* lineage in the Sierra Madre and
Medicine Bow Mountains is considered here on the basis of phylogenetic,
morphological, and ecological data. For the loci comprising individual gene
trees, reciprocal monophyly was not observed in either rITS or *rps* intron
(Figures \@ref(fig:FigureResultsBayesrITS) and
\@ref(fig:FigureResultsBayesrpsIntron)).
Phylogenetic support in the concatenated MrBayes consensus tree
(Figure \@ref(fig:FigureResultsBayesMultiLocus)) and multi-locus BEAST maximum
credibility clade tree (Figure \@ref(fig:FigureResultsBeastMultiLocus))
does support divergence of these populations relative to *P. vitulifera* of
Colorado. Modeling dispersal with contemporary geographic location similarly
supports divergence of these populations from *P. vitulifera*, with
exception to a single specimen from Hog Park in the Sierra Madre Range
(Dorn 10105, Figure \@ref(fig:FigureResultsBeastGeography)). While sampling bias
has been recently indicated as potentially influencing Brownian motion
continuous phylogeography [@Kalkauskas2021], population genetic studies of
*Calochortus gunnisonii* have revealed genetic differentiation of high-elevation
Medicine Bow and Sierra Madre populations relative to populations north of the
Wyoming Basin [@FullerMcGalughlin2019].

\paragraph{} Species delimitation under the multispecies coalescent via
\*BEAST yielded the highest support for divergence when prior *P. saximontana*
subspecies were lumped with *P. didymocarpa* (Hypotheses 3 and 4, Figure
\@ref(fig:FigureResultsBeastSpecies)). In considering species delimitation,
the degree of gene exchange (e.g., immigration) and time since divergence
provide potential measures for evolutionary independence of lineages
[@HeyPinho2012]. Statistical delimitation methods based on the multispecies
coalescent may overestimate genetic structure as species lineages, which has
implications for any method utilizing species as a unit of analysis
[@SukumaranKnowles2017]. The phylogenies reported here consist of a modest
number of specimens sampled from a relatively small number of genetic loci.
While not fully resolved, evidence for lineage separation of the disputed
*P. vitulifera* specimens as a new taxon was also observed in morphological and
ecological traits. Reviewed specimens are found generally above 2,100m, at
a higher elevation relative to nearby *P. acutifolia* and allopatric
*P. vitulifera* of Colorado (Figures \@ref(fig:FigureResultsPhysariaElevation)
and \@ref(fig:FigureDiscussionElevationSEINet)).
Globally, endemic plant species have been shown
to correlate with increases in elevation [@Steinbauer2016].

```{r FigureDiscussionElevationSEINet, fig.cap=caption$FigureDiscussionElevationSEINet, fig.scap="Colorado and south-central Wyoming elevation map"}

caption$FigureDiscussionElevationSEINet <-
  paste("Colorado and south-central Wyoming elevation map.")

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/DiscussionElevationSEINet",
    ext = knit_variables$graphics
  )
)

```

\paragraph{} Specimen occurrences obtained from SEINet for *P. floribunda*
subspecies and *P. bellii* highlight the allopatric distribution of Colorado
*Physaria* species from the Medicine Bow and Sierra Madre Range populations
\@ref(fig:FigureIntroCarbonWyo). Notably, the Front Range provides a
geographic barrier, with *P. vitulifera* and *P. bellii* limited to the
eastern slope. Similarly, the *P. floribunda* subspecies from Middle Park
to the southwest are separated from the hypothesized *Physaria* lineage by
the Bears Ears Range with populations north and northwest of North Park in
Grand County, Colorado. Populations of questioned specimens examined are
distributed from the Sierra Madre West of Encampment and East through the
Medicine Bow Range. After accounting for collection date, these specimens
show a marked reduction in average mature fruit length and width relative
to other nearby species (Figure \@ref(fig:FigureResultsPhysariaFruits)).
Traditionally, *P. floribunda* have been differentiated by recurved, nearly
pendant pedicels and pinnatifid basal leaves [@OKane2010]. Future phylogenetic
and population genetic analyses should include both *P. floribunda* subspecies
of Middle Park and *P. bellii* of the eastern slope of the Front Range.

```{r SectionSouthwesternEndemics}

Thesis::knitr_section(
  knitr_title = paste("Southwest Wyoming Endemics"),
  knitr_type = "section", chunk_type = knit_variables$pandoc
)

```

\paragraph{} In the multilocus analyses, the southwestern Wyoming endemics
*P. condensata* and *P. dornii* were consistently supported in a clade with
individuals of *P. integrifolia*. @Rollins1939 initially recognized a
west-central *Physaria* variety as distinct from northern *P. didymocarpa*
populations, with similar fruit characteristics as subsp. *didymocarpa*
(a broad replum with more than 2 ovules per locule), and transitional
characteristics of *P. acutifolia* (entire basal leaves and appressed silique
trichomes). These observations, @Rollins1983 note on the relatively inflated,
thin-walled fruits of variety *integrifolia* adapted for wind dispersal, and
additional SEM imaging formed the basis of a species description by @Lichvar1984.
The first of three condensed rosette growth populations was described by
@Rollins1939 as *P. condensata*, delineated by stems of $<$ 1cm and a rosette
diameter of $<$ 5cm. @Lichvar1983 differentiated the condensed habit *P. dornii*
from *P. condensata* by relatively enlarged basal leaves and fruits.
Traditional morphological delineations for the southwestern Wyoming endemics
are concordant with observations reported here: more than 2 ovules per locule
(Figures \@ref(fig:FigureResultsPhysariaOvules) and
\@ref(fig:FigureDiscussionOvules)),
fruit size differences (Figure \@ref(fig:FigureResultsPhysariaFruits)),
differences in growth habit (Figure \@ref(fig:FigureResultsPhysariaHabit)),
a broad replum (Figure \@ref(fig:FigureDiscussionReplumShape)),
appressed silique trichomes (Figure \@ref(fig:FigureDiscussionFruitTrichomes)),
and entire leaf margins (Figure \@ref(fig:FigureDiscussionBasalLeafMargins)).
These species are geographically separated from *P. didymocarpa* subsp.
*didymocarpa* by the Wind River and Absaroka Ranges
(Figure \@ref(fig:FigureDiscussionElevationSouthwestWY)).

\clearpage

```{r FigureDiscussionElevationSouthwestWY, fig.cap=caption$FigureDiscussionElevationSouthwestWY, fig.scap="Southwestern Wyoming elevation map"}

caption$FigureDiscussionElevationSouthwestWY <-
  paste("Southwestern Wyoming elevation map.")

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/DiscussionElevationSouthwestWY",
    ext = knit_variables$graphics
  )
)

```

\clearpage

\vspace*{-24pt}

```{r SectionPhysariaDidymocarpaSubspecies}

Thesis::knitr_section(
  knitr_title = paste("*Physaria didymocarpa* subspecies"),
  knitr_type = "section", chunk_type = knit_variables$pandoc
)

```

\paragraph{} @Rollins1984 described two new varieties of *P. saximontana*,
*saximontana* and *dentata*, with a primary distribution south of the range
of *P. didymocarpa*. These varieties were primarily differentiated from
*P. didymocarpa* on the basis of silique morphology, where mature fruits of
the *P. saximontana* varieties lack a basal sinus and the replum orifice
attaches at the base of the valve. Variety *dentata*, described from collections
in Glacier and Park Counties of Montana, was differentiated from variety
*saximontana* on the basis of toothed margins and an elongated style.
While this trait can be difficult to discern from herbarium specimens, review
of Montana *Physaria didymocarpa* specimens suggests the silique morphology
is common throughout the range of *P. didymocarpa*, putative *P. saximontana*
subspecies, and fruiting specimens of *P. eburniflora* across the Beaver Rim
Divide (Figure \@ref(fig:FigureDiscussionFruitApices)). Similarly, dentate
or otherwise toothed margins can be observed throughout the range of
*P. didymocarpa* in Montana and northwestern Wyoming
(Figure \@ref(fig:FigureDiscussionBasalLeafMargins)), though entire margins
have been observed in many Fremont County, Wyoming specimens.
The species tree analyses showed relatively higher support for lumping
*P. saximontana* subspecies with *P. didymocarpa* subspecies *didymocarpa*
(Hypotheses 3 and 4, Figure \@ref(fig:FigureResultsBeastSpecies)).
Phylogenetic analyses (Figures \@ref(fig:FigureResultsBeastMultiLocus) and
\@ref(fig:FigureResultsBeastGeography) and silique morphology suggest a possible
sister taxa relationship with *P. brassicoides*, though a gap in geographic
distribution from the southern range of *P. didymocarpa* and nominal
*P. saximontana* is apparent (Figure \@ref(fig:FigureDiscussionElevationBigHorn)).

\paragraph{} Prior to the description of *P. saximontana* varieties,
@Rollins1981 described the second condensed growth habit species as
*P. eburniflora* from specimens collected in Carbon County, Wyoming.
While flowering specimens are observed with pale-yellow petals, fruiting
specimens were described as having shallow basal sinuses differentiated from
the more-or-less equal above and below sinuses of *P. didymocarpa* and
*P. acutifolia*. Additionally, the rosette leaves of *P. eburniflora* were
described as nearly orbicular with entire margins. Fruiting specimens reviewed
here suggest growth habit (Figure \@ref(fig:FigureResultsPhysariaHabit)) and
to a degree mature fruit size (Figure \@ref(fig:FigureResultsPhysariaFruits))
may be the best characters for differentiating *P. eburniflora* from
fully resolved, species tree analyses indicate a possible sister taxa
relationship between *P. eburniflora* and *P. didymocarpa* subspecies including
*P. saximontana* (Hypothesis 4, Figure \@ref(fig:FigureResultsBeastSpecies))).
The primary range of *P. eburniflora* extends east from the Green and Ferris
Mountains to the western foothills of the Laramie Range
(Figure \@ref(fig:FigureDiscussionElevationBigHorn)).

\clearpage

```{r FigureDiscussionElevationBigHorn, fig.cap=caption$FigureDiscussionElevationBigHorn, fig.scap="\\textit{Physaria} of the Bighorn Basin and Beaver Rim Divide", fig.pos="h"}

caption$FigureDiscussionElevationBigHorn <-
  paste(
    ifelse(
      test = knit_variables$pandoc == "latex",
      yes = "\\textit{Physaria}",
      no = "*Physaria*"
    ),
    "of the Bighorn Basin and Beaver Rim Divide."
  )

knitr::include_graphics(
  path = fs::path(
    here::here(), "inst/figures/DiscussionElevationBigHorn",
    ext = knit_variables$graphics
  )
)

```


