---
title: "Project Data"
bibliography: "`r fs::path_package(package = 'thesis', 'manuscript/references.bib')`"
link-citations: true
nocite: >
    @R-readxl, @R-dplyr
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

source_url <- "https://github.com/jasonratcliff/thesis/tree/master/"
```

[//]: # (TODO <3/26/23>: Error from magrittr pipe export in thesis NAMESPACE)

```{r setup}
library(thesis)
library(magrittr) # TODO
```

# Voucher Specimens

```{r specimens}
# Assign R6 `SpecimenMap` object from voucher specimens
vouchers <- SpecimenMap$new(
    records = thesis::herbarium_specimens,
    identifier = 'prior_id'
  )
```

Herbarium specimens borrowed from lending institutions (RM, NY, MO, F, ISTC,
MONTU, MONT, RSA-POM, UC, UTC, GH, US, CAS, IDS) were compiled into the document
[`vouchers.xlsx`](`r fs::path(source_url, "data-raw/specimens/vouchers.xlsx")`).
Recorded voucher data include locality, identification history,
collector, collection number, date, institution, geographic coordinates,
elevation, ecological description, and measurements of continuous and
discrete morphological traits. In total, `r vouchers$census()[['distinct']]`
unique collections were reviewed. The R script
[`vouchers.R`](`r fs::path(source_url, "data-raw/specimens/vouchers.R")`)
formats specimen data for downstream distribution, morphological, and
phylogenetic analysis. Briefly, this script filters records by genus,
standardizes collection dates with `lubridate` [@R-lubridate],
extracts voucher annotation history, and converts locality elevation units.
Annotation synonyms are substituted according to @OKane2010.
Specimens can be accessed as an exported `tibble::tibble`
object from the package namespace.

## DNA Sampling

<details>
  <summary>Code</summary>
  <p>

```{r specimenSamples}
#| message: false
vouchers$limit(
  west = -115.2,
  east = -103,
  south = 37,
  north = 49.1
)

dna_records <- vouchers$map(
  legend = "Prior\nAnnotation",
  expand = TRUE,
  sf_states = c("New Mexico", "Arizona", "Nevada")
) +
  vouchers$repel(
    vouchers = thesis::dna_specimens,
    repel.params = list(
      na.rm = TRUE,
      alpha = 0.85,
      size = 2.75,
      xlim = c(-Inf, NA),
      ylim = c(-Inf, NA),
      segment.size = 0.5,
      segment.alpha = 1,
      arrow = grid::arrow(length = grid::unit(0.01, "npc")),
      box.padding = 0.5
    )
  )

cowplot::save_plot(
  plot = dna_records,
  file = "dna.png",
  base_height = 16, base_width = 12
)
```

  </p>
</details>

```{r fig-specimenMap}
#| echo: false
#| fig.cap: Distribution of sampled DNA specimens.
#| out.width: 100%
knitr::include_graphics(
  path = "dna.png",
  dpi = 400
)
```

# References
