---
title: "BEAST Methods"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Running BEAUti

```
java -jar /Applications/BEAST.v1.10.4/lib/beauti.jar
```

## Running BEAST

```
# Open BEAST to run .xml file
java -jar /Applications/BEAST.v1.10.4/lib/beast.jar
```


# Installations

## BEAST

- Installation [Overview](https://beast.community/installing)
	- Mac [Install](https://beast.community/install_on_mac)
	- CLI [Install](https://beast.community/install_on_unix#homebrew-package-manager-for-mac-os-x)

## Java

- For macOS, [install](http://beast.community/install_on_mac#installing-beast) recommends using Java version 8.  Needed for:
    - `beauti.jar`
    - `beast.jar`

### Java Downloads

- [Maintained Version Downloads](https://www.oracle.com/java/technologies/javase-jsp-downloads.html)
    - Java Development Kit [Version 8](https://www.oracle.com/java/technologies/jdk8-downloads.html)
        - BEAST requires Java version 6 or 8
    - Downloading a Java Runtime Enviroment (JRE) installs java as a *plug in*
        - Install the comparable Java Development Kit (JDK)

> Warning: Do not install the Java software offered on the http://java.com website. This will install java only as a plug in to your web browser. This version of Java will not be able to run BEAST.
  - Mac [BEAST Installation](https://beast.community/install_on_mac)

- From instructions for [OS X JDK 8 Installation](https://docs.oracle.com/javase/8/docs/technotes/guides/install/mac_jdk.html)

### Java Specifications

```{r javaVers, engine='bash', eval=FALSE}
java -version
```

```{r javaPath, engine='bash', eval=FALSE}
/usr/libexec/java_home -V
```

